Architectures and Open pNets

### 2018-02-11 2018-Arch-pNets/arch-pNets.tex
Added compositionality of safety properties *with proof*

\newcommand{\mdash}[1][]{---#1}
\newcommand{\ndash}{--}
\newcommand{\ie}[1][\ ]{i.e.#1}
\newcommand{\etc}[1][\ ]{etc.#1}
\newcommand{\eg}[1][\ ]{e.g.#1}
\newcommand{\cf}[1][\ ]{cf.#1}
\newcommand{\wrt}[1][\ ]{w.r.t.#1}

@@

\newcommand{\true} {\ensuremath{\mathtt{t\!t}}}
\newcommand{\false}{\ensuremath{\mathtt{f\!f}}}
\newcommand{\noop} {\ensuremath{\mathsf{skip}}}

\newcommand{\data}{\ensuremath{\sD}}
\newcommand{\guards}[1]{\ensuremath{\sB_{#1}}}
\newcommand{\exprs}[1]{\ensuremath{\sE_{#1}}}
\newcommand{\valuations}[1]{\ensuremath{\sV_{#1}}}
\newcommand{\val}[3][]{%
  \ensuremath{\sigma^{#2}_{#3}%
    \ifthenelse{\isempty{#1}}{}{(#1)}%
  }%
}
%% \DeclareMathOperator{\supp}{supp}
\newcommand{\supp}[1]{\ensuremath{\mathrm{supp}(#1)}}
\newcommand{\semopen}[1]{\ensuremath{[{#1}]}}
\newcommand{\semclosed}[1]{\ensuremath{\llbracket{#1}\rrbracket}}
\newcommand{\reachable}[1]{\ensuremath{\mathit{reachable}({#1})}}
\newcommand{\IMextend}[2]{\ensuremath{#1 \ltimes #2}}
\newcommand{\arcomp}{\oplus}
\newcommand{\Arcomp}{\Bigoplus}
\newcommand{\expmix}{\oplus}

@@

\todoSBin{Preliminaries on the universal(?) data domain $\data$
  and valuations.}

\begin{definition}[Components]
  \label{defn:component}
  A \emph{component} is a tuple $(Q, q^0, P, V, \val{0}{},
  \goesto{})$, where
  \begin{itemize}
  \item $Q$ is a set of \emph{states}, with $q^0 \in Q$ the
    \emph{initial state}, 
  \item $P$ is a set of \emph{ports},
  \item $V$ is a set of \emph{component variables},
  \item $\val{0}{} : V \rightarrow \data$ is an \emph{initial
    valuation} of the component variables, 
  \item $\goesto{}\, \subseteq
    Q \times 2^P \times \guards{V} \times \exprs{V} \times Q$
%
    is a \emph{transition relation}, with transitions
    labelled by triples consisting of an \emph{interaction}
    $a \subseteq P$, a Boolean \emph{guard} $g \in
    \guards{V}$ and an \emph{update expression} $e \in
    \exprs{V}$.
  \end{itemize}
%
  We call the pair of sets $(P,V)$ the \emph{interface} of the
  component.\,\footnote{%
%
  In practice, component variables are split in two sub-sets:
  local and exported variables.  Only exported variables belong
  to the component interface and can be used for the interactions
  with other components (see \defn{im}).  However, in the context
  of this paper we can simplify by omitting this separation.
%
  }
%
  Notations $q \goesto{a} q'$, $q \goesto{a}$ and $q
  \notgoesto{a}$ are as usual; for a component $B$, we denote
  $Q_B$, $q^0_B$, $P_B$, $V_B$, $\val{0}{B}$ and $\goesto[B]{}$
  the constituents of $B$.
\end{definition}

\begin{definition}[Component semantics]
  \label{defn:comp:semantics}
  The \emph{open semantics} of a component $B = (Q, q^0, P, V,
  \val{0}{}, \goesto{})$ is given by the LTS $\semopen{B} = (S,
  s^0, \goesto{})$, where $S = Q \times \valuations{V}$, $s^0 =
  (q^0, \val{0}{})$ and $\goesto{}$ is the minimal transition
  relation satisfying the following rule:
  %
  \begin{equation}
    \label{eq:comp:semantics}
    \infer{
      q \goesto{a, g, e} q'
      \and
      v \models g
      \and
      v' = v[e]
    }{
      (q, v) \goesto{a} (q', v')
    }
    \,.
  \end{equation}
  The \emph{closed semantics} of $B$ is given by the LTS
  $\semclosed{B} \bydef{=}\reachable{\semopen{B}}$, comprising
  only the reachable states of $\semopen{B}$.
\end{definition}

\begin{definition}[Interaction model \& composition]
  \label{defn:im}
  Let $\cB = \{B_1,\dots,B_n\}$ be a finite set of
  components with $B_i = (Q_i, q^0_i, P_i, V_i, \val{0}{i},
  \goesto{})$,\footnote{%
%
    Here and below, we skip the index on the transition
    relation $\goesto{}$, since it is always clear from the
    context.
%
} such that all their respective components (\ie $Q_i$, $P_i$,
  and $V_i$) are pairwise disjoint, \ie $\forall i \neq j,\ Q_i
  \cap Q_j = P_i \cap P_j = V_i \cap V_j = \emptyset$.  Let $P =
  \bigcup_{i = 1}^n P_i$ and $V = \bigcup_{i = 1}^n V_i$.

  An \emph{interaction model over $(P,V)$} is a set $\Gamma
  \subseteq 2^P \times \guards{V} \times \exprs{V}$, such that,
  for any $(a, g, e) \in \Gamma$, the guard and update expression
  associated to the \emph{interaction} $a$ satisfy, respectively,
  $g \in \guards{V_a}$ and $e \in \exprs{V_a}$, with $V_a
  \bydef{=} \bigcup_{i:\, a \cap P_i \neq \emptyset} V_i$.  We
  denote $\supp{a} \bydef{=} \setdef{i \in [1,n]}{a \cap P_i \neq
    \emptyset}$ the \emph{support} of the interaction $a$.
\noteLH{why not introduce supp earlier and use it for $V_a$?}
  %% We call the set of ports $P$ the \emph{domain} of the
  %% interaction model.

  The \emph{composition of $\cB$ with the interaction model
    $\Gamma$} is the component $\Gamma(\cB) = (Q, q^0, P, V,
  \val{0}, \goesto{})$, where $Q = \prod_{i=1}^n Q_i$, $q^0 =
  q_1^0\dots q_n^0$, $\val{0}{}: V \rightarrow \data$ is such
  that, for any $v \in V_i$, $\val[v]{0}{} = \val[v]{0}{i}$, and
  $\goesto{}$ is the minimal transition relation satisfying
  following rules:
%
  \begin{mathpar}
    \infer{
      q_i \goesto{\emptyset, \true, \noop} q_i'
    }{
      q_1 \dots q_i \dots q_n \goesto{\emptyset, \true, \noop} q_1 \dots q_i' \dots q_n
    }\,,

    \infer{    
      (a, g, e) \in \Gamma
      \and
      \forall i \in \supp{a}, q_i \goesto{a \cap P_i, g_i, e_i} q_i'
      \and
      \forall i \not\in \supp{a}, q_i = q_i'
      \\\\
      G = g \land \bigwedge_{i \in \supp{a}} g_i
      \and
      E = e; (e_i)_{i \in \supp{a}}
    }{
      q_1\dots q_n \goesto{a, G, E} q_1'\dots q_n'
    }\,.
  \end{mathpar}
\end{definition}

@@

\begin{definition}[Architecture]
  \label{defn:arch}
  An \emph{architecture} is a tuple $A = (\cC, P_A, V_A, \Gamma)$,
  where $\cC$ is a finite set of \emph{coordinating components}
  with pairwise disjoint sets of ports and variables, such that
  $\bigcup_{C \in \cC} P_C \subseteq P_A$ and
  $\bigcup_{C \in \cC} V_C \subseteq V_A$, and
  $\Gamma \subseteq 2^{P_A} \times \guards{V_A} \times \exprs{V_A}$
  is an interaction model over $(P_A, V_A)$.
\end{definition}

\begin{definition}[Application of an architecture]
  \label{defn:arch:application}
  Let $A = (\cC, P_A, V_A, \Gamma)$ be an architecture and let $\cB$
  be a set of components, such that
%
  \begin{align}
    \bigcup_{B \in \cB} P_B \cap \bigcup_{C \in \cC} P_C = \emptyset\,,
    &&
    P_A \subseteq P \bydef{=} \bigcup_{B \in \cB \cup \cC} P_B\,,
    \\
    \bigcup_{B \in \cB} V_B \cap \bigcup_{C \in \cC} V_C = \emptyset\,,
    &&
    V_A \subseteq V \bydef{=} \bigcup_{B \in \cB \cup \cC} V_B\,,
  \end{align}
%
  and, denoting, for any $(a, g, e) \in \Gamma$, $V_a \bydef{=}
  \bigcup_{B \in \supp{a}} V_B$, we have $g \in \guards{V_a}$ and
  $e \in \exprs{V_a}$.
%
  The \emph{application of the architecture $A$ to the set of
  components $\cB$} is the component $ A(\cB) \bydef{=}
  (\IMextend{\Gamma}{P})(\cC \cup \cB)$, where
%
  \begin{equation}
    \label{eq:im:extension}
    \IMextend{\Gamma}{P} \bydef{=}
    \setdefb{
      (a \cup a', g, e)
    }{
      (a, g, e) \in \Gamma, a' \subseteq P \setminus P_A
    }
  \end{equation}
%
  denotes the \emph{extension} of the interaction model $\Gamma$
  to the set of ports $P$.
\end{definition}

An architecture $A$ enforces coordination constraints on the
components in $\cB$.  The interface $(P_A, V_A)$ of an
architecture $A$ contains all ports of the coordinating
components $\cC$ and some additional ports, which must belong to
the components in $\cB$.  In the application $A(\cB)$, the ports
belonging to $P_A$ can only participate in the interactions
defined by the interaction model $\Gamma$ of $A$.  Ports which do
not belong to $P_A$ are not restricted and can participate in any
interaction.  In particular, they can join the interactions in
$\Gamma$ (see \eq{im:extension}).  If the interface of the
architecture covers all ports of the system, \ie $P = P_A$, we
have $P\setminus P_A = \emptyset$ and the only interactions
allowed in $A(\cB)$ are those belonging to $\Gamma$.  Notice also
that the restrictions imposed by \defn{arch:application} on the
set of operand components $\cB$ ensure that an interaction in
$\IMextend{\Gamma}{P}$ can only refer to variables of the
participating components.
%
Finally, the definition of $\IMextend{\Gamma}{P}$ requires that
an interaction from $\Gamma$ be involved in every interaction
belonging to $\IMextend{\Gamma}{P}$.  To allow the ports from $P
\setminus P_A$ to be fired independently in $A(\cB)$, one must
have $(\emptyset, \true, \noop) \in \Gamma$.

\begin{definition}[Composition of architectures]
  \label{defn:arch:composition}
  Let $A_i = (\cC_i, P_{A_i}, V_{A_i}, \Gamma_i)$, for $i = 1,2$,
  be two architectures.  The \emph{composition} of $A_1$ and
  $A_2$ is the architecture $A_1 \arcomp A_2 = (\cC_1 \cup \cC_2,
  P_{A_1} \cup P_{A_2}, V_{A_1} \cup V_{A_2}, \Gamma)$, where
%
  \begin{equation}
    \label{eq:arch:composition}
    \Gamma = \setdefb{
      (a, g_1 \land g_2, e_1 \expmix e_2) 
    }{
      (a \cap P_{A_i}, g_i, e_i) \in \Gamma_i,
      \text{ for } i = 1,2
    }
    \,.
  \end{equation}
\end{definition}

\begin{proposition}[Properties of $\arcomp$]
  \label{prop:arcomp:nice}
  Architecture composition $\arcomp$ is commutative and
  associative; it is idempotent if all coordinating components
  are deterministic; $A_{id} = \bigl(\emptyset, \emptyset,
  \emptyset, \{\emptyset\}\bigr)$ is its neutral element, \ie for
  any architecture $A$, we have $A \oplus A_{id} = A$.
  Furthermore, for any component $B$, we have $A_{id}(B) = B$.
  %PCA: replaced ``holds'' by ``we have''
\end{proposition}
%
\begin{proof}[Sketch of the proof]
  Commutativity and associativity follow from the corresponding
  properties of set union, Boolean conjunction and
  \todoSB{Check ``semi''}\hl{semilattice meet}.
%
  Suppose we have two architectures $A = A'$.  This does not
  necessarily mean that their sets of coordinating components
  coincide.  However, if all the involved coordinating components
  are deterministic, then, in any state of $(A \arcomp A')(\cB)$,
  both architectures will impose the same restrictions, enabling
  the same interactions between the coordinating and operand
  components.  Hence, we have $(A \arcomp A')(\cB) = A(\cB) =
  A'(\cB)$.  Since this holds for any set of components $\cB$, we
  conclude that $A \arcomp A' = A = A'$.
%
  The properties of $A_{id}$ follow immediately from the
  definitions of architecture application and composition.
\end{proof}

@@

>>>>

\newcommand{\mdash}[1][]{---#1}
\newcommand{\ndash}{--}
\newcommand{\ie}[1][\ ]{i.e.#1}
\newcommand{\etc}[1][\ ]{etc.#1}
\newcommand{\eg}[1][\ ]{e.g.#1}
\newcommand{\cf}[1][\ ]{cf.#1}
\newcommand{\wrt}[1][\ ]{w.r.t.#1}
\newcommand{\resp}[1][\ ]{resp.#1}

@@

\newcommand{\true} {\ensuremath{\mathtt{t\!t}}}
\newcommand{\false}{\ensuremath{\mathtt{f\!f}}}
\newcommand{\noop} {\ensuremath{\emptyset}} % \mathsf{skip}

\newcommand{\data}{\ensuremath{\sD}}
\newcommand{\guards}[1]{\ensuremath{\sB_{#1}}}
\newcommand{\exprs}[1]{\ensuremath{\sE_{#1}}}
\newcommand{\valuations}[1]{\ensuremath{\sV_{#1}}}
\newcommand{\val}[3][]{%
  \ensuremath{#1{\sigma}^{#2}_{#3}}%
}
\newcommand{\primeit}[1]{#1'}
\newcommand{\doubleprimeit}[1]{#1''}

\newcommand{\export}[1][]{\ensuremath{\varepsilon_{#1}}}
\newcommand{\valdiff}[2]{\ensuremath{#1 \triangle #2}}
\newcommand{\supp}[1]{\ensuremath{\mathrm{supp}(#1)}}
\newcommand{\semopen}[1]{\ensuremath{[{#1}]}}
\newcommand{\semclosed}[1]{\ensuremath{\llbracket{#1}\rrbracket}}
\newcommand{\reachable}[1]{\ensuremath{\mathit{reachable}({#1})}}
\newcommand{\IMextend}[2]{\ensuremath{#1 \ltimes #2}}
\newcommand{\arcomp}{\oplus}
\newcommand{\arequiv}{\equiv}
\newcommand{\Arcomp}{\Bigoplus}
\newcommand{\expmix}{\wedge}
\newcommand{\order}{\leqslant}

\makeatletter
\newcommand{\doubletilde}[1]{{%
  \mathpalette\double@tilde{#1}%
}}
\newcommand{\double@tilde}[2]{%
  \sbox\z@{$\m@th#1\tilde{#2}$}%
  \ht\z@=.9\ht\z@
  \tilde{\box\z@}%
}
\makeatother

\newcounter{tempctr}
\newcommand{\breakenumistart}{%
  \setcounter{tempctr}{\value{enumi}}%
  \end{enumerate}%
}
\newcommand{\breakenumiend}{%
  \begin{enumerate}%
  \setcounter{enumi}{\value{tempctr}}%
}

@@

\todoSBin{Preliminaries on the universal(?) data domain $\data$
  and valuations.

For two valuations $\val{1}{}, \val{2}{} : V \rightarrow \data$,
we denote \[\valdiff{\val{1}{}}{\val{2}{}} \bydef{=} \setdefb{v
  \in V}{\val{1}{}(v) \neq \val{2}{}(v)}\] the set of variables
that are assigned different values by the two valuations.

To guarantee preservation of properties guards and expressions
have to be \emph{monotonic}.  In the case of a trivial order, \ie
when nothing is comparable, \emph{any guard or expression is
  monotonic}.  Thus, this is not really a constraint!
}

\begin{definition}[Components]
  \label{defn:component}
  A \emph{component} is a tuple $(Q, q^0, V, \val{0}{}, P,
  \export{}, \goesto{})$, where
  \begin{itemize}
  \item $Q$ is a set of \emph{states}, with $q^0 \in Q$ the
    \emph{initial state}, 
  \item $V$ is a set of \emph{component variables},
  \item $\val{0}{} : V \rightarrow \data$ is an \emph{initial
    valuation} of the component variables, 
  \item $P$ is a set of \emph{ports}, with
    \noteSB{Added this}\hl{$\export : P \rightarrow 2^V$ 
    the \emph{export function},}
  \item $\goesto{}\, \subseteq
    Q \times 2^P \times \guards{V} \times \exprs{V} \times Q$
%
    is a \emph{transition relation}, with transitions
    labelled by triples consisting of an \emph{interaction}
    $\emptyset \neq a \subseteq P$, a Boolean \emph{guard} $g \in
    \guards{V}$ and an \emph{update expression} $e \in
    \exprs{V}$.
  \end{itemize}
%
  We call \noteSB{Maybe rather $(P, \export)$}\hl{the pair of sets $(V,P)$} the \emph{interface} of the
  component.\,\footnote{%
%
  In practice, component variables are split in two sub-sets:
  local and exported variables.  Only exported variables belong
  to the component interface and can be used for the interactions
  with other components (see \defn{im}).  However, in the context
  of this paper we can simplify by omitting this separation.
%
  }
%
  Notations $q \goesto{a} q'$ and $q \goesto{a}$ are as usual;
  for a component $B$, we denote $Q_B$, $q^0_B$, $P_B$, $V_B$ and
  $\val{0}{B}$ the corresponding constituents of $B$.  We will
  skip the index on the transition relations $\goesto{}$, since it
  is always clear from the context.
\end{definition}

\begin{definition}[Component semantics]
  \label{defn:comp:semantics}
  The \emph{open semantics} of a component $B = (Q, q^0, V,
  \val{0}{}, P, \export, \goesto{})$ is given by the LTS $\semopen{B} = (S,
  s^0, \goesto{})$, where $S = Q \times \valuations{V}$, $s^0 =
  (q^0, \val{0}{})$ and $\goesto{}$ is the minimal transition
  relation satisfying the rule
  %
  \begin{equation}
    \label{eq:comp:semantics}
    \infer{
      q \goesto{a, g, e} q'
      \and
      \val{}{} \models g
      \and
      \val[\primeit]{}{} = \tilde{\val{}{}}[e]
      \and
      \valdiff{\val{}{}}{\tilde{\val{}{}}} \subseteq \export(a)
    }{
      (q, \val{}{}) \goesto{a,\tilde{\val{}{}}} (q', \val[\primeit]{}{})
    }
    \,.
  \end{equation}
  %
  \todoSB{Do we actually need this?}
  \hl{The \emph{closed semantics} of $B$ is given by the LTS
  $\semclosed{B} \bydef{=}\reachable{\semopen{B}}$, comprising
  only the reachable states of $\semopen{B}$.}
\end{definition}

The use of the intermediate valuation $\tilde{\val{}{}}$ in the
conclusion and the third premise of rule \eq{comp:semantics}
allows some of the variables to get new values before the
transition is actually fired.  Thus the component is \emph{open}
to the exchange of data with its environment.  However, the
fourth premise in \eq{comp:semantics} restricts the set of
variables, which can be affected by such data transfer, to those
that are exported through the ports participating in the
interaction.

\begin{definition}[Interaction model \& composition]
  \label{defn:im}
  Consider a finite set of components $\cB = (Q_i, q^0_i, V_i,
  \val{0}{i}, P_i, \export[i], \goesto{})^{i \in I}$, such that
  all their respective components (\ie $Q_i$, $P_i$, and $V_i$)
  are pairwise disjoint, \ie $\forall i \neq j,\ Q_i \cap Q_j =
  P_i \cap P_j = V_i \cap V_j = \emptyset$.  Let $P = \bigcup_{i
  \in I} P_i$ and $V = \bigcup_{i \in I} V_i$.
%
  For a set of ports $a \subseteq P$, we denote $\supp{a}
  \bydef{=} \setdef{i \in I}{a \cap P_i \neq \emptyset}$ the
  \emph{support} of $a$,
  \noteSB{Updated this}\hl{and $V_a \bydef{=} \bigcup_{i \in
    \supp{a}} \export[i](a \cap P_i)$ the set of component
  variables exported for $a$.}

  An \emph{interaction model over \hl{$(V,P)$}} is a set $\Gamma
  \subseteq 2^P \times \guards{V} \times \exprs{V}$, such that,
  for any $(a, g, e) \in \Gamma$, the guard and update expression
  associated to the \emph{interaction} $a$ satisfy, respectively,
  $g \in \guards{V_a}$ and $e \in \exprs{V_a}$.\footnote{%
%
    Notice that this definition allows $(\emptyset, \true,
    \noop)$ and $(\emptyset, \false, \noop)$ to be included in
    $\Gamma$.
%
  }
  %% We call the set of ports $P$ the \emph{domain} of the
  %% interaction model.

  The \emph{composition of $\cB$ with the interaction model
    $\Gamma$} is the component
  $\Gamma(\cB) = (Q, q^0, V, \val{0}{}, P, \export, \goesto{})$,
  where
%
  $Q = \prod_{i \in I} Q_i$;
%
  $q^0 = (q_i^0)^{i \in I}$;
%
  $\val{0}{}: V \rightarrow \data$ is such that, for any $v \in
  V_i$, $\val{0}{}(v) = \val{0}{i}(v)$;
%
  $\export : P \rightarrow 2^V$ is such that, for any $p \in
  P_i$, $\export(p) = \export[i](p)$;
%
  and $\goesto{}$ is the minimal transition relation satisfying
  the rule
%
  \begin{gather}
%%     \label{eq:im:empty}
%%     \infer{
%%       q_j \goesto{\emptyset, g, e} q_j'
%%       \and
%%       \forall i \neq j, q_i = q_i'
%%     }{
%%       (q_i)^{i \in I} \goesto{\emptyset, g, e} (q_i')^{i \in I}
%%     }\,,
%% %
%%     \\[0.5\baselineskip]
    \label{eq:im:int}
%
    \infer{    
      (a, g, e) \in \Gamma
      \and
      a \neq \emptyset
      \and
      \forall i \in \supp{a}, q_i \goesto{a \cap P_i, g_i, e_i} q_i'
      \and
      \forall i \not\in \supp{a}, q_i = q_i'
%
      \\\\
      \textstyle
%
      G = g \land \bigwedge_{i \in \supp{a}} g_i
      \and
      E = e; (e_i)^{i \in \supp{a}}
    }{
      (q_i)^{i \in I} \goesto{a, G, E} (q_i')^{i \in I}
    }\,.
  \end{gather}
\end{definition}

@@

\begin{definition}[Architecture]
  \label{defn:arch}
  An \emph{architecture} is a tuple $A = (\cC, V_A, P_A, \Gamma)$,
  where $\cC$ is a finite set of \emph{coordinating components}
  with pairwise disjoint sets of ports and variables, such that
  $\bigcup_{C \in \cC} P_C \subseteq P_A$ and
  $\bigcup_{C \in \cC} V_C \subseteq V_A$, and
  $\Gamma \subseteq 2^{P_A} \times \guards{V_A} \times \exprs{V_A}$
  is an interaction model over \hl{$(V_A, P_A)$}.
\end{definition}

\begin{definition}[Application of an architecture]
  \label{defn:arch:application}
  Let $A = (\cC, V_A, P_A, \Gamma)$ be an architecture and let $\cB$
  be a set of components, such that
%
  \begin{align}
    \bigcup_{B \in \cB} V_B \cap \bigcup_{C \in \cC} V_C = \emptyset\,,
    &&
    V_A \subseteq V \bydef{=} \bigcup_{B \in \cB \cup \cC} V_B\,,
    \\
    \bigcup_{B \in \cB} P_B \cap \bigcup_{C \in \cC} P_C = \emptyset\,,
    &&
    P_A \subseteq P \bydef{=} \bigcup_{B \in \cB \cup \cC} P_B\,,
  \end{align}
%
  and
  %% , denoting, for any $(a, g, e) \in \Gamma$, $V_a \bydef{=}
  %% \bigcup_{B \in \supp{a}} \export[B](a \cap P_B)$, we have
  $g \in \guards{V_a}$ and
  $e \in \exprs{V_a}$ (see \defn{im} for the notation $V_a$).
%
  The \emph{application of the architecture $A$ to the set of
  components $\cB$} is the component $ A(\cB) \bydef{=}
  (\IMextend{\Gamma}{P})(\cC \cup \cB)$, where
%
  \begin{equation}
    \label{eq:im:extension}
    \IMextend{\Gamma}{P} \bydef{=}
    \setdefb{
      (a \cup a', g, e)
    }{
      (a, g, e) \in \Gamma, a' \subseteq P \setminus P_A
    }
  \end{equation}
%
  denotes the \emph{extension} of the interaction model $\Gamma$
  to the set of ports $P$.
\end{definition}

\todoSBin{Define equivalence $\arequiv$?}

An architecture $A$ enforces coordination constraints on the
components in $\cB$.  The interface \hl{$(V_A, P_A)$} of an
architecture $A$ contains all ports of the coordinating
components $\cC$ and some additional ports, which must belong to
the components in $\cB$.  In the application $A(\cB)$, the ports
belonging to $P_A$ can only participate in the interactions
defined by the interaction model $\Gamma$ of $A$.  Ports which do
not belong to $P_A$ are not restricted and can participate in any
interaction.  In particular, they can join the interactions in
$\Gamma$ (see \eq{im:extension}).  If the interface of the
architecture covers all ports of the system, \ie $P = P_A$, we
have $P\setminus P_A = \emptyset$ and the only interactions
allowed in $A(\cB)$ are those belonging to $\Gamma$.  Notice also
that the restrictions imposed by \defn{arch:application} on the
set of operand components $\cB$ ensure that an interaction in
$\IMextend{\Gamma}{P}$ can only refer to variables of the
participating components.
%
Finally, the definition of $\IMextend{\Gamma}{P}$ requires that
an interaction from $\Gamma$ be involved in every interaction
belonging to $\IMextend{\Gamma}{P}$.  To allow the ports from $P
\setminus P_A$ to be fired independently in $A(\cB)$, one must
have $(\emptyset, \true, \noop) \in \Gamma$.

\begin{definition}[Composition of architectures]
  \label{defn:arch:composition}
  Let $A_i = (\cC_i, V_{A_i}, P_{A_i}, \Gamma_i)$, for $i = 1,2$,
  be two architectures.  The \emph{composition} of $A_1$ and
  $A_2$ is the architecture $A_1 \arcomp A_2 = (\cC_1 \cup \cC_2,
  V_{A_1} \cup V_{A_2}, P_{A_1} \cup P_{A_2}, \Gamma)$, where
%
  \begin{equation}
    \label{eq:arch:composition}
    \Gamma = \setdefb{
      (a, g_1 \land g_2, e_1 \expmix e_2) 
    }{
      (a \cap P_{A_i}, g_i, e_i) \in \Gamma_i,
      \text{ for } i = 1,2
    }
    \,.
  \end{equation}
\end{definition}

\begin{proposition}[Properties of $\arcomp$]
  \label{prop:arcomp:nice}
  Architecture composition $\arcomp$ is commutative and
  associative; it is idempotent if all coordinating components
  are deterministic; $A_{id} = \bigl(\emptyset, \emptyset,
  \emptyset, \{\emptyset\}\bigr)$ is its neutral element, \ie for
  any architecture $A$, we have $A \oplus A_{id} \arequiv A$.
  Furthermore, for any component $B$, we have $A_{id}(B) = B$.
\end{proposition}
%
\begin{proof}[Sketch of the proof]
  Commutativity and associativity follow from the corresponding
  properties of set union, Boolean conjunction and
  \todoSB{Check ``semi''}\hl{semilattice meet}.
%
  Suppose we have two architectures $A = A'$.  This does not
  necessarily mean that their sets of coordinating components
  coincide.  However, if all the involved coordinating components
  are deterministic, then, in any state of $(A \arcomp A')(\cB)$,
  both architectures will impose the same restrictions, enabling
  the same interactions between the coordinating and operand
  components.  Hence, we have $(A \arcomp A')(\cB) = A(\cB) =
  A'(\cB)$.  Since this holds for any set of components $\cB$, we
  conclude that $A \arcomp A' \arequiv A \arequiv A'$.
%
  The properties of $A_{id}$ follow immediately from the
  definitions of architecture application and composition.
\end{proof}

@@

For a component $B$, we denote $S_{\semopen{B}}$,
$s^0_{\semopen{B}}$ the corresponding constituents of
$\semopen{B}$.

\begin{definition}[Safety properties]
  \label{defn:property}
  Let $B$ be a component.  A \emph{safety property} (below,
  simply \emph{property}) of $B$ is a state predicate $\Phi:
  S_{\semopen{B}} \rightarrow \sB$, such that $\bigl((q,
  \val{}{}) \models \Phi\bigr) \land (\val[\primeit]{}{} \order
  \val{}{})$ implies $(q, \val[\primeit]{}{}) \models \Phi$,
  where we write $(q, \val{}{}) \models \Phi$ iff $\Phi(q,
  \val{}{}) = \true$.  A property $\Phi$ is \emph{initial} if
  $s^0_{\semopen{B}} \models \Phi$\todoSB{Do we need this?}\hl{;
    it is \emph{reachable} iff there exists a possibly empty path
    $s^0_{\semopen{B}} \goesto{a^1, \val{1}{}} s^1 \goesto{a^2,
      \val{2}{}} \cdots \goesto{a^n, \val{n}{}} s^n$, such that
    $s^n \models \Phi$}.
\end{definition}

\todoSB{Cite the original paper somewhere.}
The main idea of our approach is that an architecture enforces
its characteristic property on the set of its operand components.
From this point of view, the set of coordinating components is
not relevant, neither are their states.  Thus, to talk about
properties enforced by architectures, we consider properties on
the unrestricted composition of the operand components as
formalized by the following definition.

\begin{definition}[Enforcing properties]
  \label{defn:impose}
  Let $A = (\cC, P_A, V_A, \Gamma)$ be an architecture; let $\cB$
  be a set of components and $\Phi$ an initial property of
  their parallel composition $A_{id}(\cB)$ (see
  \prop{arcomp:nice}).  We say that \emph{$A$ enforces $\Phi$ on
    $\cB$} iff, for every state $s = (s_b, s_c)$ reachable in
  $\semopen{A(\cB)}$, with $s_b \in \prod_{B\in \cB}
  S_{\semopen{B}}$ and $s_c \in \prod_{C \in \cC}
  S_{\semopen{C}}$, we have $s_b \models \Phi$.
\end{definition}

According to the above definition, when we say that an
architecture enforces some property $\Phi$, it is implicitly
assumed that $\Phi$ is initial for the coordinated components.
Below, we omit mentioning this explicitly.

\begin{theorem}[Preserving enforced properties]
  \label{thm:combining}
  Let $\cB$ be a set of components; let $A_i = (\cC_i, V_{A_i},
  P_{A_i}, \Gamma_i)$, for $i = 1,2$, be two architectures
  enforcing on $\cB$ the properties $\Phi_1$ and $\Phi_2$
  respectively.  The composition $A_1 \arcomp A_2$ enforces on
  $\cB$ the property $\Phi_1 \land \Phi_2$.
\end{theorem}

\begin{lemma}
  \label{lem:onlyone}
  Let $A = (\cC, V_A, P_A, \Gamma)$ be an architecture and denote
  by $\Gamma_\cC \bydef{=}
%
  \setdef{
    (a \cap P_\cC, \true, \noop)
  }{
    (a, g, e) \in \Gamma
  }$, with $P_\cC = \bigcup_{C \in \cC} P_C$,
%  
  the projection of $\Gamma$ onto the coordinating components of
  $A$.  Consider the architecture $A' = (\{C'\}, V_A, P_A,
  \Gamma)$, where $C' = \Gamma_\cC(\cC)$.  For any set of
  components $\cB$, satisfying the conditions of
  \defn{arch:application}, we have
  $\semopen{A(\cB)} = \semopen{A'(\cB)}$.
\end{lemma}
%
\begin{proof}
  First of all, notice that, by \defn{im},
  $V_{C'} = \bigcup_{C \in \cC} V_C$ and $P_{C'} = P_\cC$.
  The conditions of \defn{arch} are satisfied and $A'$ is indeed
  an architecture.  Furthermore, $\cB$ satisfies the conditions
  of \defn{arch:application} \wrt $A'$.  Hence, the component
  $A'(\cB)$ is well defined.

  Clearly the state spaces, initial states and interfaces of both
  components coincide.  Thus we only have to prove that so do the
  transition relations.  Let us assume that
  $\cC = (C_i)^{i \in I}$ and $\cB = (B_j)^{j \in J}$.
  We will use $q^i, q^{i\prime}$ to denote the states of
  $C_i$ and $q_j, q_j'$ to denote the states of $B_j$,
  and similarly for the valuations of variables.

  By \defn{comp:semantics}, a transition
%
  \begin{equation}
    \label{eq:lem1:trans:sem}
    (q^i, \val{i}{})^{i \in I} (q_j, \val{}{j})^{j \in J}
    \goesto{a, \val[\tilde]{}{}}
    (q^{i\prime}, \val{i\prime}{})^{i \in I}
    (q_j', \val{\prime}{j})^{j \in J}
  \end{equation}
%  
  is a transition in $\semopen{A(\cB)}$ (\resp
  $\semopen{A'(\cB)}$) iff
%
  \begin{equation}
    \label{eq:lem1:trans}
    (q^i)^{i \in I} (q_j)^{j \in J}
    \goesto{a, G, E}
    (q^{i\prime})^{i \in I} (q_j')^{j \in J}
    \,,
  \end{equation}
%
  is a transition in $A(\cB)$ (\resp $A'(\cB)$) and
%
  \begin{mathpar}
    (\val{i}{})^{i \in I} (\val{}{j})^{j \in J}
    \models G
    \,,
    \and
    (\val{i\prime}{})^{i \in I} (\val{\prime}{j})^{j \in J}
    = \val[\tilde]{}{}[E]
    \,,
    \and
    \valdiff{(\val{i}{})^{i \in I} (\val{}{j})^{j \in J}}{
      \val[\tilde]{}{}} \subseteq \export(a)
    \,.   
  \end{mathpar}

  By \defn{im}, \eq{lem1:trans} is a transition in $A(\cB)$
  iff $a \neq \emptyset$, $(a, g, e) \in \IMextend{\Gamma}{P}$,
  where $P = \bigcup_{B \in \cB \cup \cC} P_B$, and
  %
  \begin{enumerate}
  \item for $i \in I$,
    $q^i \goesto{a \cap P_{C_i}, g_{C_i}, e_{C_i}} q^{i\prime}$
    is a transition in $C_i$, or
    $a \cap P_{C_i} = \emptyset$ and $q^i = q^{i\prime}$,
    $g_{C_i} = \true$ and $e_{C_i} = \noop$\,;
  \item for $j \in J$,
    $q_j \goesto{a \cap P_{B_j}, g_{B_j}, e_{B_j}} q_j'$
    is a transition in $B_j$, or
    $a \cap P_{B_j} = \emptyset$ and $q_j = q_j'$,
    $g_{B_j} = \true$ and $e_{B_j} = \noop$\,;
  \item $G = g \land \bigwedge_{i \in I} g_{C_i} \land
    \bigwedge_{j \in J} g_{B_j}$ and
    $E = e; (e_{C_i})^{i \in I} (e_{B_j})^{j \in J}$.
  \end{enumerate}
  %
